---
title: "DDoS Attack Mitigation"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---
```{r setup, echo=FALSE, include=FALSE}

source("ddosm.R")

```

# Overview

## Parameters

### Dataset

We used a dataset with 2^24 packets in the detection phase.

### Sensitivity Coefficients

Our experiments with tcad_m_levels.py show us candidate values for k, as follows: 

Log2(m) | k     | FPR
--------|-------|-----
14      | 4.125 | 1.7%
16      | 4.500 | 1.6%
18      | 3.625 | 0.0%

### TCAD Traces

[Source: DDoS Mitigation.ipynb, section Finding TCAD Values]

  tcad_m_2_14_k_4.125.log
  tcad_m_2_16_k_4.500.log
  tcad_m_2_18_k_3.625.log

# Workload Characterization

## Reading TCAD Trace Files

```{r read_tcad_traces, include=TRUE}

tcad_m14 = read_tcad_trace(tcad_m14_trace)
tcad_m16 = read_tcad_trace(tcad_m16_trace)
tcad_m18 = read_tcad_trace(tcad_m18_trace)

```

## Entropy Overview

```{r entropy_overview, include=TRUE}

title = "Entropy for each Observation Window "

get_plot_tcad(tcad_m14, tcad_m14_k) + labs(title=str_c(title,"(m = 2^14)"))
get_plot_tcad(tcad_m16, tcad_m16_k) + labs(title=str_c(title,"(m = 2^16)"))
get_plot_tcad(tcad_m18, tcad_m18_k) + labs(title=str_c(title,"(m = 2^18)"))

graph_path = "graphs"

ggsave(path=graph_path, filename="entropy_all.svg")

```

## Entropy Under Attack

We have the graph for the entire experiments. Now we need to focus in the attack.

```{r entropy_under_attack, include=TRUE}

title = "Entropy for each Observation Window - Attack Phase "

get_plot_tcad_attack(tcad_m14, tcad_m14_k, log2n, 14) + labs(title=str_c(title,"(m = 2^14)"))
get_plot_tcad_attack(tcad_m16, tcad_m16_k, log2n, 16) + labs(title=str_c(title,"(m = 2^16)"))
get_plot_tcad_attack(tcad_m18, tcad_m18_k, log2n, 18) + labs(title=str_c(title,"(m = 2^18)"))
ggsave(path=graph_path, filename="entropy_attack.svg")


```

# Attack Mitigation

## Loading

```{r}

log2m = as.integer(14)

if (log2m == 14)  {
  pcap_csv = pcap_csv_m14
} else if (log2m == 16) {
  pcap_csv = pcap_csv_m16
} else if (log2m == 18) {
  pcap_csv = pcap_csv_m18
} else {
  stop("Invalid log2m value.")
}

attack_first_ow = attack_first(log2n, log2m) + 2 # Diversion begins at the second OW under attack. 
attack_last_ow = attack_last(log2n, log2m)

packets = read_pcap_csv(log2n, log2m, pcap_csv) %>% filter(ow >= attack_first_ow, ow <= attack_last_ow)


```

## Typical Deltas

For each OW, what are the typical frequency deltas for attack packets? 

```{r typical_deltas, include=TRUE}

deltas = summarize_deltas(log2n, log2m, packets)

dot_size = 2.0

graph_path = str_c("graphs_", log2m)

delta_plot_options = list(
  geom_point(aes(y=diffq1), color="blue4", size=dot_size),
  geom_point(aes(y=diffq2), color="yellow4", size=dot_size),
  geom_point(aes(y=diffq3), color="orangered4", size=dot_size))

deltas %>% 
  ggplot(aes(x=ow, shape=attack)) + delta_plot_options

ggsave(path=graph_path, filename="diffs.svg")

# It would be cool to plot several box plots. 

```


## Classification and Statistics

We want to find the best value for t, considering its effect on the accuracy of the mechanism. 

Base Stats and Confidence Intervals: TPR and FPR are proportions whose values we need to estimate.

Let X be a random variable whose value is 1 when a classification succeeds, and 0 otherwise. Consequently, X follows a a Bernoulli distribution with probability p. We can use our sample to estimate the mean value of p. This sample mean has a sampling error approximately equal to sqrt(p * (1-p) / n), where n is the sample size (the total number of classification events).     

1. Considering the attack as a whole, what are the TPR and the FPR? 
2. Considering OWs, what are the TPR and the FPR? 
3. Considering OWs, what is the 95% confidence interval for the mean TPR and FPR?  

```{r fn_stats, include=TRUE}

# summary_all = tibble() 

# for(diff_threshold in seq(-2048, 2048, by=128)) {
# 
#   divert = function(diff) (diff >= diff_threshold)
#   packets = packets %>% mutate(diverted=divert(diff)) %>% group_by(ow)
#   stats = get_stats(packets)
#   summary = get_summary(stats, log2n, log2m, diff_threshold)
#   summary_all = bind_rows(summary_all, summary)
# 
# }

summary_all = read_csv2("summary_all.csv")

summary_all

```

## Graphs

It is interesting to observe what happens **over time**, OW after OW. 

```{r classification_graphs, include=TRUE}

# Log2m == 14: seq(-2048, 2048, by=128) ; ideal = 512
# Log2m == 16: seq(-2048, 2048, by=128) ; ideal = 768
# Log2m == 18: seq(-2048, 2048, by=128) ; ideal = 1280

if (log2m == 14)  {
  diff_threshold = 512
} else if (log2m == 16) {
  diff_threshold = 768
} else if (log2m == 18) {
  diff_threshold = 1280
} else {
  stop("Invalid log2m value.")
}

divert = function(diff) (diff >= diff_threshold)
packets = packets %>% mutate(diverted=divert(diff)) %>% group_by(ow)

graph_marked_good(packets)
graph_marked_evil(packets)
graph_false_negatives(packets)
ggsave(path=graph_path, filename="false_negatives.svg")
graph_false_positives(packets)
ggsave(path=graph_path, filename="false_positives.svg")
graph_results(packets)
ggsave(path=graph_path, filename="results.svg")

```

Adjust scale, add legitimate and attack percentage lines. 

## Address Counts

```{r frequencies}

# graph_options = list(geom_line(), scale_color_manual(values=c("seagreen4", "orangered1")))
# 
# src_distinct = packets %>% group_by(ow, attack, diff) %>% summarize(srcs=n_distinct(src))
# src_distinct %>% ggplot(aes(diff, srcs, color=attack)) + graph_options
# 
# ggsave(path=graph_path, filename="src_delta_distribution.svg")
# 
# dst_distinct = packets %>% group_by(ow, attack, diff) %>% summarize(dsts=n_distinct(dst))
# dst_distinct %>% ggplot(aes(diff, dsts, color=attack)) + graph_options
# 
# ggsave(path=graph_path, filename="dst_delta_distribution.svg")

```
