---
title: "DDoS Attack Mitigation"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---
```{r setup, echo=FALSE, include=FALSE}

source("ddosm.R")

knitr::opts_chunk$set(tidy=TRUE, tidy.opts=list(width.cutoff=160))

```

# Overview

## Parameters

We used a packet trace with n=2^24 packets in the detection phase.

```{r parameters}

log2n = as.integer(24)  # n is passed to trafg. 
n = as.integer(2^log2n)

```

## Packet Traces 

```{r loading}

pcap_dir = "~/p4sec/ddosm-p4/pcaps"

pcap_csv_m14 = str_c(pcap_dir,"/ddos20m14b/if3_attack_out.csv")
pcap_csv_m16 = str_c(pcap_dir,"/ddos20m16b/if3_attack_out.csv")
pcap_csv_m18 = str_c(pcap_dir,"/ddos20m18b/if3_attack_out.csv")

log2m = as.integer(14)
packets_m14 = read_pcap_csv(log2n, log2m, pcap_csv=pcap_csv_m14) 
attack_first_ow = attack_first(log2n, log2m) + 2 # Diversion begins at the second OW under attack. 
attack_last_ow = attack_last(log2n, log2m)
packets_m14 = packets_m14 %>% filter(ow >= attack_first_ow, ow <= attack_last_ow) 

log2m = as.integer(16)
packets_m16 = read_pcap_csv(log2n, log2m, pcap_csv=pcap_csv_m16)  
attack_first_ow = attack_first(log2n, log2m) + 2 # Diversion begins at the second OW under attack. 
attack_last_ow = attack_last(log2n, log2m)
packets_m16 = packets_m16 %>% filter(ow >= attack_first_ow, ow <= attack_last_ow) 

log2m = as.integer(18)
packets_m18 = read_pcap_csv(log2n, log2m, pcap_csv=pcap_csv_m18) 
attack_first_ow = attack_first(log2n, log2m) + 2 # Diversion begins at the second OW under attack. 
attack_last_ow = attack_last(log2n, log2m)
packets_m18 = packets_m18 %>% filter(ow >= attack_first_ow, ow <= attack_last_ow) 

```

## TCAD Traces

Sensitivity Coefficients 

Our experiments with tcad_m_levels.py show us candidate values for k, as follows: 

Log2(m) | k     | FPR
--------|-------|-----
14      | 4.125 | 1.7%
16      | 4.500 | 1.6%
18      | 3.625 | 0.0%

[Source: DDoS Mitigation.ipynb, section Finding TCAD Values]

```{r }

trace_dir = "~/p4sec/ddosm-p4/lab/ddos20/tcad_logs"

tcad_m14_k = 4.125
tcad_m16_k = 4.500
tcad_m18_k = 3.625

tcad_m14_trace = str_c(trace_dir,"/tcad_m_2_14_k_4.125.log")
tcad_m16_trace = str_c(trace_dir,"/tcad_m_2_16_k_4.500.log")
tcad_m18_trace = str_c(trace_dir,"/tcad_m_2_18_k_3.625.log")

tcad_m14 = read_tcad_trace(tcad_m14_trace)
tcad_m16 = read_tcad_trace(tcad_m16_trace)
tcad_m18 = read_tcad_trace(tcad_m18_trace)



```

# Workload Characterization

## Entropy Overview

```{r entropy_overview, include=TRUE}

title = "Entropy for each Observation Window "

get_plot_tcad(tcad_m14, tcad_m14_k) + labs(title=str_c(title,"(m = 2^14)"))
ggsave(path="graphs_14", filename="entropy_all.svg")

get_plot_tcad(tcad_m16, tcad_m16_k) + labs(title=str_c(title,"(m = 2^16)"))
ggsave(path="graphs_16", filename="entropy_all.svg")

get_plot_tcad(tcad_m18, tcad_m18_k) + labs(title=str_c(title,"(m = 2^18)"))
ggsave(path="graphs_18", filename="entropy_all.svg")

```

## Entropy Under Attack

We have the graph for the entire experiments. Now we need to focus in the attack.

```{r entropy_under_attack, include=TRUE}

title = "Entropy for each Observation Window - Attack Phase "

get_plot_tcad_attack(tcad_m14, tcad_m14_k, log2n, 14) + labs(title=str_c(title,"(m = 2^14)"))
ggsave(path="graphs_14", filename="entropy_attack.svg")

get_plot_tcad_attack(tcad_m16, tcad_m16_k, log2n, 16) + labs(title=str_c(title,"(m = 2^16)"))
ggsave(path="graphs_16", filename="entropy_attack.svg")

get_plot_tcad_attack(tcad_m18, tcad_m18_k, log2n, 18) + labs(title=str_c(title,"(m = 2^18)"))
ggsave(path="graphs_18", filename="entropy_attack.svg")

```

# Attack Mitigation

## Typical Deltas

For each OW, what are the typical frequency deltas for attack packets? 

```{r typical_deltas}

for(log2m in seq(14, 18, by=2)) {

  log2m = as.integer(log2m)
  
  if (log2m == 14)  {
    packets = packets_m14 
  } else if (log2m == 16) {
    packets = packets_m16 
  } else if (log2m == 18) {
    packets = packets_m18 
  } else {
    stop("Invalid log2m value.")
  }
  
  deltas = summarize_deltas(log2n, log2m, packets)

  graph_path = str_c("graphs_", log2m)
  m_annotation = str_c("(m = 2^", log2m, ")") 
  dot_size = 2.0

  delta_plot_options = list(
#    geom_point(aes(y=diffq1), color="blue4", size=dot_size),
    geom_point(aes(y=diffq2), color="blue4", size=dot_size),
#    geom_point(aes(y=diffq3), color="orangered4", size=dot_size),
    labs(x="Observation Window", y="Median Address Frequency Variation", title = str_c("Median Address Frequency Variations ", m_annotation)),
    theme_classic())
  
  deltas %>% ggplot(aes(x=ow, shape=attack)) + delta_plot_options
  
  ggsave(path=graph_path, filename="diffs.svg")  
  
}


```

## Diversion Thresholds

We want to find the best value for t, considering its effect on the accuracy of the mechanism. 

Base Stats and Confidence Intervals: TPR and FPR are proportions whose values we need to estimate. Let X be a random variable whose value is 1 when a classification succeeds, and 0 otherwise. Consequently, X follows a a Bernoulli distribution with probability p. We can use our sample to estimate the mean value of p. This sample mean has a sampling error approximately equal to sqrt(p * (1-p) / n), where n is the sample size. Considering the attack as a whole, what is the 95% confidence interval for the mean TPR and FPR?       

```{r best_threshold, include=TRUE}

summary_all = tibble() 

for(log2m in seq(14, 18, by=2)) {

  log2m = as.integer(log2m)
  
  if (log2m == 14)  {
    packets = packets_m14 
  } else if (log2m == 16) {
    packets = packets_m16 
  } else if (log2m == 18) {
    packets = packets_m18 
  } else {
    stop("Invalid log2m value.")
  }

  for(diff_threshold in seq(-2048, 2048, by=128)) {
  
    divert = function(diff) (diff >= diff_threshold)
    packets = packets %>% mutate(diverted=divert(diff)) %>% group_by(ow)
    stats = get_stats(packets)
    summary = get_summary(stats, log2n, log2m, diff_threshold)
    summary_all = bind_rows(summary_all, summary)
  
  }
  

}

summary_all

# I have saved a good summary in this file: 

# summary_all = read_csv2("summary_all.csv")

# summary_all

# Log2m == 14: ideal = 512
# Log2m == 16: ideal = 768
# Log2m == 18: ideal = 1280


```

## Diversion Statistics

```{r diversion_stats}

stats_all = tibble()

for(log2m in seq(14, 18, by=2)) {

  log2m = as.integer(log2m)
  
  if (log2m == 14)  {
    packets = packets_m14 
    diff_threshold = 512
  } else if (log2m == 16) {
    packets = packets_m16 
    diff_threshold = 768
  } else if (log2m == 18) {
    packets = packets_m18 
    diff_threshold = 1280
  } else {
    stop("Invalid log2m value.")
  }

  divert = function(diff) (diff >= diff_threshold)
  packets = packets %>% mutate(diverted=divert(diff)) %>% group_by(ow)

  stats = get_stats(packets) %>% mutate(log2n=log2n, log2m=log2m, t=diff_threshold)
  stats_all = bind_rows(stats_all, stats)
    
  graph_path = str_c("graphs_", log2m)
  m_annotation = str_c("(m = 2^", log2m, ")")   
  
  # It would be cool to plot several box plots. 
  graph_false_negatives(packets)
  ggsave(path=graph_path, filename="false_negatives.svg")
  graph_false_positives(packets)
  ggsave(path=graph_path, filename="false_positives.svg")
  graph_results(packets, m_annotation)
  ggsave(path=graph_path, filename="results.svg")

}

```

## Address Counts

```{r frequencies}

# graph_options = list(geom_line(), scale_color_manual(values=c("seagreen4", "orangered1")))
# 
# src_distinct = packets %>% group_by(ow, attack, diff) %>% summarize(srcs=n_distinct(src))
# src_distinct %>% ggplot(aes(diff, srcs, color=attack)) + graph_options
# 
# ggsave(path=graph_path, filename="src_delta_distribution.svg")
# 
# dst_distinct = packets %>% group_by(ow, attack, diff) %>% summarize(dsts=n_distinct(dst))
# dst_distinct %>% ggplot(aes(diff, dsts, color=attack)) + graph_options
# 
# ggsave(path=graph_path, filename="dst_delta_distribution.svg")

```
