---
title: "DDoS Attack Mitigation (Short Runs)"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---
```{r setup, echo=FALSE, include=FALSE}

source("ddosm.R")

knitr::opts_chunk$set(tidy=TRUE, tidy.opts=list(width.cutoff=160))

```

# Overview

## Parameters

We used a packet trace with n=2^24 packets in the detection phase.

```{r parameters}

log2n = as.integer(24)   

```

## Packet Traces 

```{r loading}

#pcap_dir = "~/p4sec/ddosm-p4/pcaps"
pcap_dir = "D:/P4/ddosm-p4/pcaps"

pcap_csv_m14 = str_c(pcap_dir,"/n_2_24_m_2_14/if3_attack_out.csv.gz")
pcap_csv_m16 = str_c(pcap_dir,"/n_2_24_m_2_16/if3_attack_out.csv.gz")
pcap_csv_m18 = str_c(pcap_dir,"/n_2_24_m_2_18/if3_attack_out.csv.gz")

packets_m14 = read_pcap_csv(log2n=log2n, log2m=14, pcap_csv=pcap_csv_m14, attack_only=FALSE) 
packets_m16 = read_pcap_csv(log2n=log2n, log2m=16, pcap_csv=pcap_csv_m16, attack_only=FALSE) 
packets_m18 = read_pcap_csv(log2n=log2n, log2m=18, pcap_csv=pcap_csv_m18, attack_only=FALSE) 

```

# Attack Mitigation

## Typical Deltas

For each OW, what are the typical frequency deltas for attack packets? 

```{r typical_deltas}

for(log2m in seq(14, 18, by=2)) { 

  log2m = as.integer(log2m)
  
  if (log2m == 14)  {
    packets = packets_m14 
  } else if (log2m == 16) {
    packets = packets_m16 
  } else if (log2m == 18) {
    packets = packets_m18 
  } else {
    stop("Invalid log2m value.")
  }
  
  deltas = summarize_deltas(log2n, log2m, packets)

  graph_path = str_c("graphics/m", log2m)
  m_annotation = str_c("(m = 2^", log2m, ")") 
  dot_size = 2.0

  delta_plot_options = list(
#    geom_point(aes(y=diffq1), color="blue4", size=dot_size),
    geom_point(aes(y=diffq2), color="blue4", size=dot_size),
#    geom_point(aes(y=diffq3), color="orangered4", size=dot_size),
    labs(x="Observation Window",
         y="Median Address Frequency Variation", 
         title = str_c("Median Address Frequency Variations ", m_annotation)),
    theme_classic())
  
  deltas %>% ggplot(aes(x=ow, shape=attack)) + delta_plot_options
  
  my_ggsave(plot=last_plot(), path=graph_path, filename="diffs")  
  
}


```

## Diversion Thresholds

We want to find the best value for t, considering its effect on the accuracy of the mechanism. 

Base Stats and Confidence Intervals: TPR and FPR are proportions whose values we need to estimate. Let X be a random variable whose value is 1 when a classification succeeds, and 0 otherwise. Consequently, X follows a a Bernoulli distribution with probability p. We can use our sample to estimate the mean value of p. This sample mean has a sampling error approximately equal to sqrt(p * (1-p) / n), where n is the sample size. Considering the attack as a whole, what is the 95% confidence interval for the mean TPR and FPR?       

Note:
For m=2^18:
The attack begins at OW #49 (attack_first()). 
Mitigation starts at OW #50, but the accuracy is unacceptably low. 
From OW #50 on, accuracy stabilizes. 


```{r best_threshold, include=TRUE}

summary_all = tibble() 

for(log2m in seq(14, 18, by=2)) { 

  log2m = as.integer(log2m)
  
  if (log2m == 14)  {
    packets = packets_m14 
    sequence = seq(-256, 256, by=16)
  } else if (log2m == 16) {
    packets = packets_m16 
    sequence = seq(-512, 512, by=32)
  } else if (log2m == 18) {
    packets = packets_m18
    sequence = seq(-1024, 1024, by=64)
  } else {
    stop("Invalid log2m value.")
  }

  for(diff_threshold in sequence) {
  
    divert = function(diff) (diff >= diff_threshold)
    packets = packets %>% 
      filter(ow >= attack_first(log2n,log2m)+2) %>%   # See note about this.
      mutate(diverted=divert(diff)) %>% 
      mutate(tp =  attack & diverted, fp = !attack & diverted) %>%
      ungroup()
    stats = get_stats(packets)
    summary = get_summary(stats, log2n, log2m, diff_threshold)
    summary_all = bind_rows(summary_all, summary)
    
  }
  
}

summary_all

# write.csv(summary_all,'data/summary_all_short.csv')

# summary_all = read.csv('data/summary_all_short.csv')


```



```{r}


summary_all %>% 
  ggplot(aes(x=t/2^(log2m/2+1), color=as.factor(log2m))) + 
  geom_line(aes(y=p_tp)) + geom_point(aes(y=p_tp), shape=5) + 
  geom_line(aes(y=p_fp)) + geom_point(aes(y=p_fp), shape=4) + 
  scale_color_manual(name="log2m", breaks=c(14,16,18), values=c("red","green","blue")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x="Normalized Mitigation Threshold (t/2^(log2(m)/2+1))", y="") + 
  theme_classic() 

graph_path = "graphics"
my_ggsave(plot=last_plot(), path=graph_path, filename="tpr_fpr")

summary_all_tpr = summary_all %>% transmute(log2m=log2m, t=t, type="TPR", mean=p_tp, lb=i95_tp_lb, ub=i95_tp_ub) 
summary_all_fpr = summary_all %>% transmute(log2m=log2m, t=t, type="FPR", mean=p_fp, lb=i95_fp_lb, ub=i95_fp_ub) 
summary_all_x = bind_rows(summary_all_tpr, summary_all_fpr)

summary_all_x

summary_all_x %>% 
  ggplot(aes(x=t/2^(log2m/2+1), color=type, shape=type)) + 
  geom_line(aes(y=mean), linetype="dashed" ) + 
  geom_point(aes(y=mean), size=2) + 
  geom_errorbar(aes(ymin=lb, ymax=ub)) + 
  scale_color_manual(name="type", values=c("red","blue")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x="Normalized Mitigation Threshold (t/2^(log2(m)/2+1))", y="") + 
  facet_grid(rows = vars(log2m), labeller = label_both) +
  theme_classic() +
  theme(legend.title = element_blank())

my_ggsave(plot=last_plot(), path=graph_path, filename="tpr_fpr_2")



```

## Diversion Statistics

```{r diversion_stats}

stats_all = tibble()

for(log2m in seq(14, 18, by=2)) {

  log2m = as.integer(log2m)
  
  if (log2m == 14)  {
    packets = packets_m14 
    diff_threshold = 128
  } else if (log2m == 16) {
    packets = packets_m16 
    diff_threshold = 256
  } else if (log2m == 18) {
    packets = packets_m18 
    diff_threshold = 512
  } else {
    stop("Invalid log2m value.")
  }

  divert = function(diff) (diff >= diff_threshold)
  packets = packets %>% 
    mutate(diverted=divert(diff)) %>%
    mutate(tp =  attack & diverted, fp = !attack & diverted) %>%
    group_by(ow)

  stats = get_stats(packets) %>% mutate(log2n=log2n, log2m=log2m, t=diff_threshold)
  stats_all = bind_rows(stats_all, stats)
    
  graph_path = str_c("graphics/m", log2m)
  m_annotation = str_c("(n = 2^27, m = 2^", log2m, ")")   
  
  # It would be cool to plot several box plots. 
  graph_false_negatives(packets)
  my_ggsave(plot=last_plot(), path=graph_path, filename="false_negatives")
  graph_false_positives(packets)
  my_ggsave(plot=last_plot(), path=graph_path, filename="false_positives")
  graph_results(packets, m_annotation)
  my_ggsave(plot=last_plot(), path=graph_path, filename="results")

}

# write.csv(stats_all,'data/stats_all_short.csv')

```

