---
title: "TCAD (Long Runs)"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---
```{r setup, echo=FALSE, include=FALSE}

source("analysis.R")

knitr::opts_chunk$set(tidy=TRUE, tidy.opts=list(width.cutoff=160))

```

# Overview

## Parameters

We used a packet trace with approximately n=2^27 (exactly n=2^17*1000) packets in the detection phase.

```{r parameters}

log2n = as.integer(27)

```

## TCAD Traces

Sensitivity Coefficients 

Our experiments with tcad_m_levels.py show us candidate values for k, as follows: 

Log2(m) | k     | FPR
--------|-------|-----
14      | 4.125 | 1.7%
16      | 4.500 | 1.6%
18      | 3.625 | 0.0%

[Source: DDoS Mitigation.ipynb, section Finding TCAD Values]

```{r }

# tcad.trace.dir = "~/p4sec/ddosm-p4/labs/ddos20_long/tcad_logs"
tcad.trace.dir = "D:/P4/ddosm-p4/labs/ddos20_long/tcad_logs"

tcad.m14.k = 4.875
tcad.m16.k = 4.875
tcad.m18.k = 3.625

tcad.m14.trace = str_c(tcad.trace.dir,"/tcad_m_2_14_k_4.875.log")
tcad.m16.trace = str_c(tcad.trace.dir,"/tcad_m_2_16_k_4.875.log")
tcad.m18.trace = str_c(tcad.trace.dir,"/tcad_m_2_18_k_3.625.log")

tcad.m14 = read_tcad_trace(tcad.m14.trace)
tcad.m16 = read_tcad_trace(tcad.m16.trace)
tcad.m18 = read_tcad_trace(tcad.m18.trace)


```

# Workload Characterization

## Entropy Overview

```{r entropy_overview, include=TRUE}

entropy.title = "Entropy for each Observation Window "

get_plot_tcad(tcad.m14, tcad.m14.k) + labs(title=str_c(entropy.title,"(m = 2^14)"))
my_ggsave(plot=last_plot(), path="graphics_m14", filename="entropy_all_n_2_27")

get_plot_tcad(tcad.m16, tcad.m16.k) + labs(title=str_c(entropy.title,"(m = 2^16)"))
my_ggsave(plot=last_plot(), path="graphics_m16", filename="entropy_all_n_2_27")

get_plot_tcad(tcad.m18, tcad.m18.k) + labs(title=str_c(entropy.title,"(m = 2^18)"))
my_ggsave(plot=last_plot(), path="graphics_m18", filename="entropy_all_n_2_27")

```

## Entropy Under Attack

We have the graph for the entire experiments. Now we need to focus in the attack.

```{r entropy_under_attack, include=TRUE}

title = "Entropy for each Observation Window - Attack Phase "

get_plot_tcad_attack(tcad.m14, tcad.m14.k, log2n, 14) + labs(title=str_c(title,"(m = 2^14)"))
my_ggsave(plot=last_plot(), path="graphics_m14", filename="entropy_attack_n_2_27")

get_plot_tcad_attack(tcad.m16, tcad.m16.k, log2n, 16) + labs(title=str_c(title,"(m = 2^16)"))
my_ggsave(plot=last_plot(), path="graphics_m16", filename="entropy_attack_n_2_27")

get_plot_tcad_attack(tcad.m18, tcad.m18.k, log2n, 18) + labs(title=str_c(title,"(m = 2^18)"))
my_ggsave(plot=last_plot(), path="graphics_m18", filename="entropy_attack_n_2_27")

```

